# LOOCV-Transformed-Kernel-Prediction
```{r setup-transformed-kernel-prediction, message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("tidyverse", "data.table", "dtplyr", "ggthemes", "viridis",
               "stringr", "forcats")

knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```



```{r Load-LOOCV-Data, cache = TRUE}
log_file <- fread("./data/derived/pred_log.txt")
ped_log <- log_file %>%
  filter(Pred1 %in% c("ped100", "ped77", "ped42"),
         Pred2 == "none",
         Pred3 == "none")
log_file <- log_file %>%
  filter(Transformation == TRUE) %>%
  rbind(., ped_log)

pred_df <- log_file %>%
  .$Job_ID %>%
  as.list() %>%
  map(~paste0("./data/derived/predictions/", ., ".RDS")) %>%
  map(readRDS) %>%
  bind_rows() %>%
  left_join(y = log_file %>% select(Job_ID, Pred1, Pred2, Pred3, Iter),
            by = "Job_ID") %>%
  unite(col = Predictor, Pred1, Pred2, Pred3, sep = "-") %>%
  mutate(Trait = fct_recode(Trait,
    "DMY" = "GTM",
    "DMC" = "GTS",
    "FAT" = "FETT",
    "PRO" = "RPR",
    "SUG" = "XZ"
  )) %>%
  mutate(Trait = fct_relevel(Trait,
                             "DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG"),
         Predictor = as.factor(Predictor)) %>%
  select(-Job_ID, -Iter)
```



## Distribution of predicted and observed values
```{r Distribution-y-yhat, cache = TRUE, fig.height = 12, fig.cap = "Frequency polygon of observed and predicted values for seven agronomic traits (columns) and nine sets of predictors (rows)."}
# Specify a unified predictor order.
pred_order <- c("ped42-none-none", "snp42-none-none", "mrna42-none-none",
                "ped100-none-none", "snp100-none-none", "ped100-snp77-none",
                "ped100-mrna42-none", "snp100-mrna42-none",
                "ped100-snp77-mrna42")
legend_labs <- expression(y, hat(y))
pred_df %>%
  filter(Predictor != "snp77-none-none") %>%
  gather(key = Value, value = `Phenotypic Value`, y, yhat) %>%
  mutate(Predictor = fct_relevel(Predictor, pred_order),
         Value = as.factor(Value)) %>%
  ggplot(aes(x = `Phenotypic Value`, color = Value)) +
  geom_freqpoly() +
  scale_color_manual(values = c("#1f78b4", "#e31a1c"), labels = legend_labs) +
  facet_grid(Predictor ~ Trait, scales = "free") +
  theme_bw() +
  theme(legend.position = "top",
        strip.text.y = element_text(size = 7)) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  ylab("Frequency")
```

## Coefficient of variation for predicted values
We are interested in the variation in the predicted values relative to the 
arithmetic mean of the predicted values, separately for each agronomic trait and
each set of predictors:

$$
\begin{equation}
  \label{eq:CV}
CV = \sqrt{\frac{\sum_{i = 1}^{N} \sigma^{2}_{\hat{y}(i)} / N}{\sum_{i=1}^{N}\hat{y}_{i} / N}},
\end{equation}
$$

where $N$ denotes the number of hybrids.

```{r Coefficient-of-Variation, cache = TRUE}
coefficient_of_variation <- function(x, y) {
  sqrt(mean(x)) / mean(y)
}
cv_df <- pred_df %>%
  split(list(.$Trait, .$Predictor)) %>%
  map(~coefficient_of_variation(x = .$var_yhat, y = .$yhat)) %>%
  stack() %>%
  separate(col = ind, into = c("Trait", "Predictor"), sep = "[.]") %>%
  rename(CV = values) %>%
  tbl_dt() %>%
  mutate(Predictor = as.factor(Predictor),
         Trait = as.factor(Trait))

overview_df <- pred_df %>%
  group_by(Predictor, Trait) %>%
  summarize(r = cor(y, yhat)) %>%
  ungroup() %>%
  as.data.frame() %>%
  left_join(y = cv_df, by = c("Predictor", "Trait")) %>%
  as_tibble() %>%
  mutate(Predictor = as.factor(Predictor),
         Predictor = forcats::fct_reorder(Predictor, CV),
         Trait = fct_relevel(Trait,
                             "DMY", "DMC", "ADF", "FAT", "PRO",
                             "STA", "SUG"),
         Group = ifelse(Predictor %in% c("snp42-none-none",
                                         "ped42-none-none",
                                         "mrna42-none-none"),
                        yes = "Reduced", no = "Full")
  ) %>%
  mutate(Group = as.factor(Group))
```



```{r Coefficient-of-Variation-Plot, cache = TRUE, fig.cap = "Coefficient of variation for seven agronomic traits and nine sets of predictors."}
overview_df %>%
  filter(Predictor != "snp77-none-none") %>%
  mutate(Predictor = fct_relevel(Predictor, pred_order)) %>%
  ggplot(aes(x = CV, y = Predictor)) +
  geom_segment(aes(yend = Predictor), xend = 0, color = "gray50") +
  geom_point(aes(color = Predictor), size = 3) +
  scale_color_viridis(discrete = TRUE, option = "inferno") +
  facet_wrap(~ Trait, scales = "free_x") +
  theme_bw() +
  theme(panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = c(0.7, 0.13)
  ) +
  guides(color = guide_legend(ncol = 2,
                              reverse = FALSE)) +
  xlab("Coefficient of Variation") +
  ylab("Predictor Set") +
  xlim(0, NA)
```

## Predictive Abilities
```{r Predictive-Ability-Plots, cache = TRUE, fig.height = 10, fig.cap = "Predictive ability for seven agronomic traits (rows) and nine predictor sets (columns). The predictor sets are split into the categories 'Reduced' (i.e. 685 hybrids) and 'Full' (i.e. 1,521 hybrids)."}
overview_df %>%
  as_tibble() %>%
  filter(Predictor != "snp77-none-none") %>%
  mutate(Predictor = fct_relevel(Predictor, pred_order),
         Group = fct_relevel(Group, "Reduced", "Full")
  ) %>%
  ggplot(aes(x = Predictor, y = r)) +
  geom_bar(stat = "identity", aes(fill = Predictor), color = "black") +
  geom_text(aes(label = round(r, digits = 3)), vjust = -0.2) +
  scale_fill_viridis(discrete = TRUE, option = "inferno") +
  facet_grid(Trait ~ Group, space = "free", scales = "free_x") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "bottom",
    legend.key = element_rect(color = "white")
  ) +
  guides(fill = guide_legend(nrow = 3,
                             byrow = TRUE,
                             keywidth = 0.25,
                             keyheight = 0.25,
                             default.unit = "inch",
                             title.position = "top")
  ) +
  ylim(c(0, 1))
```

